In this thesis, we compared two approaches to regression with density responses: the
functional regression model after transformation to the space of log quantile
densities \parencites{PetersenMüller2016}{KokoszkaEtAl2019}, and global Fréchet
regression in the Wasserstein space \parencite{PetersenMüller2019}.

The simulation study makes the practical trade-off between the two methods concrete.
When the true regression function is linear in the space of quantile functions and
the densities have differing supports (Section~\ref{sec:app_dgp}), Fréchet
regression is both correctly specified and unburdened by nuisance parameters: its
mean ISE falls from $0.18$ to $0.04$ as the sample size grows from 50 to 200, while
the LQD method remains stuck above $31$. The additional scenarios in
Section~\ref{sec:app_additional_scenarios} disentangle why. With a common support,
the very same comparison shrinks from a factor of several hundred to well within an
order of magnitude, showing that almost all of the LQD method's error is caused by
having to re-estimate the support location that the transformation discards --- not
by the regression in the transformed space itself. Conversely, when the truth is
linear in LQD space, the LQD method wins, but the misspecified Fréchet estimator
still settles at a small bias of about $4 \times 10^{-3}$; when neither model holds,
the Fréchet method degrades far more gracefully (mean ISE of about 7 against about
31). Finally, an estimation step for the densities themselves dominates the error of
both methods and erodes their convergence, underlining the importance of the
bandwidth choice discussed in Section~\ref{sec:app_results}.

Beyond the numbers, the two methods differ markedly in their ease of application.
Fréchet regression adapts to the geometry of the Wasserstein space directly, needs
no transformation and back-transformation, produces valid distributions by
construction, and required no tuning in our implementation beyond the monotonicity
projection of Algorithm~\ref{alg:quadprog}. The transformation method is more
delicate: it requires the choice of an anchor point and an estimator for it,
inherits the numerical fragility of quantile densities near the boundaries
(Appendix~\ref{sec:numerics}), and its results are only as good as the linearity
assumption in the transformed space. Its strength is the converse of that
assumption: when a functional linear model in LQD space is plausible --- and the
supports are common or anchored --- it attains the better rate, and it gives access
to the full fda toolbox in the transformed space.

Several extensions suggest themselves. Local Fréchet regression or
Nadaraya--Watson-type smoothing would address the bias that both global methods
exhibit under the nonlinear DGP of Section~\ref{sec:app_additional_scenarios}.
Richer families of transport maps \parencite{PanaretosZemel2016} could generate more
demanding shape variation than our sine-perturbation scenario. For inference rather
than point estimation, \textcite{PetersenLiuDivani2021} provide confidence bands for
Fréchet regression in the Wasserstein context. And on the practical side, a
sample-size-dependent bandwidth choice in the density estimation step, together with
a more sophisticated estimator of the support location than linear interpolation,
would give the comparison in the estimated-density scenarios sharper contours.

In summary, when densities are the response, global Fréchet regression is the more
robust default: it is simpler to apply, never leaves the density space, and its
losses under misspecification are modest, whereas the transformation method's losses
under support uncertainty are not.
